---
title: "Comparisons take 1"
author: "S. Haire"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(readr)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(ggcorset)
```

```{r dataprep, include=FALSE}
d<-read_csv('./data/1996_2015_Neo_trop_RE2_BA_TPH_dataformat.csv')
## fixing names and factor levels
d$severity<-recode(d$severity, H="High", M="Moderate", L="Low", U="Unburned")
d$severity<-factor(d$severity, levels=c("High", "Moderate", "Low", "Unburned")) 
d$BROAD_CT<-recode(d$BROAD_CT, "mixed conifer Forest" = "Mixed Conifer",
                   "Ponderosa pine forest" = "Ponderosa Pine",
                   "Broadleaf Evergreen Woodland" = "Broadleaf Evergreen")
d$year=factor(d$year, levels = c("year 1996", "year 2015"))
d$year<-recode(d$year, "year 2015" = "2015", "year 1996" = "1996")

```
### A first report exploring data distributions between sample years, burn severity classes, and vegetation/cover types.

### Basic boxplots and comparison statistics
The first plot compares the distribution of basal area values measured in 1996 and 2015. The p-value gives a sense of difference in the two groups from a t-test. The graphs divide the data into cover types (columns doing down) and severity levels (rows going across).

```{r compareba, echo=FALSE, warning=FALSE, fig.height = 9, fig.width = 9, fig.align = "center"}
my_comparisons = list(c("2015", "1996")) # compare by sample year
p<-ggboxplot(d, x = "year", y = "BA", color = "BROAD_CT", palette = "jco", 
  facet.by=c("severity", "BROAD_CT"), add="point") + 
  stat_compare_means(comparisons = my_comparisons, method="t.test", label.y=c(60)) +
  theme_gray() + theme(strip.text.x = element_text(size = 10, colour = "black")) +
  theme(strip.text.y = element_text(size = 10, colour = "black"))
ggpar(p, legend.title="", legend="top", xlab="Year",ylab="Basal Area")

```

The second plots compare measured values of Trees per Hectare.

```{r comparetph, echo=FALSE, warning=FALSE, fig.height = 9, fig.width = 9, fig.align = "center"}
my_comparisons = list(c("2015", "1996")) # compare by sample year
p<-ggboxplot(d, x = "year", y = "TPH", color = "BROAD_CT", palette = "jco", 
  facet.by=c("severity", "BROAD_CT"), add="point") + 
  stat_compare_means(comparisons = my_comparisons, method="t.test", label.y=c(1400)) +
  theme_gray() + theme(strip.text.x = element_text(size = 10, colour = "black")) +
  theme(strip.text.y = element_text(size = 10, colour = "black"))
ggpar(p, legend.title="", legend="top", xlab="Year",ylab="Trees per Hectare")

```
### Trying out a new package!
Coming soon..."corset plots." These are pretty cool:) 
https://cran.r-project.org/web/packages/ggcorset/vignettes/corset_plot_intro.html

