---
title: "corset_test"
author: "S. Haire"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
## issues with installation following change to windows 11...
library(readr)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(reshape2)
#library(MetBrewer)
#Rtools::install_github("kbelisar/ggcorset") # not working w Rtools or devtools
library(ggcorset)
library(patchwork) # installed from command line
```

```{r dataprep, include=FALSE}
d<-read_csv('./data/1996_2015_Neo_trop_RE2_BA_TPH_dataformat.csv')
## facet by cover types and severity
d$severity<-recode(d$severity, H="High", M="Moderate", L="Low", U="Unburned")
d$severity<-factor(d$severity, levels=c("High", "Moderate", "Low", "Unburned")) 
d$BROAD_CT<-recode(d$BROAD_CT, "mixed conifer Forest" = "Mixed Conifer Forest",
                   "Ponderosa pine forest" = "Ponderosa Pine Forest")
d$year2<-recode(d$year, "year 2015" = "2015", "year 1996" = "1996")
d2<- d %>% select(Plot_ID, Area, BA, TPH, severity, BROAD_CT, year2)
## reaarange the data
yr1<-d2[d2$year2=="1996",]
yr2<-d2[d2$year2=="2015",]
yr_df<-full_join(yr1, yr2, by="Plot_ID")
names(yr_df)<-c("Plot_ID", "Island1", "BA.time1", "TPH.time1", "severity.time1",
      "covertype.time1", "time1", "Island2", "BA.time2", "TPH.time2", "severity.time2",
      "covertype.time2", "time2")
##d3<-d2 %>% group_by(Plot_ID) %>% summarize(badiff=diff(BA), tphdiff=diff(TPH)) # 285
## calc diff in BA
yr_df$badiff=yr_df$BA.time2 - yr_df$BA.time1
# Direction of Change for BA
yr_df$direction <- ifelse(yr_df$badiff<0,"Decrease",
                              ifelse(yr_df$badiff>0,"Increase","No Change"))
yr_df$direction <- factor(yr_df$direction, 
                              levels = c("Increase","No Change","Decrease"))
## one cover type at a time, due to limitations of facet() in ggcorset
ct1<-yr_df%>% filter(covertype.time1=="Ponderosa Pine Forest")
ct2<-yr_df%>% filter(covertype.time1=="Mixed Conifer Forest")
ct3<-yr_df%>% filter(covertype.time1=="Conifer Woodland")
ct4<-yr_df%>% filter(covertype.time1=="Broadleaf Evergreen Woodland")
ct5<-yr_df%>% filter(covertype.time1=="Transition Forest")
ct6<-yr_df%>% filter(covertype.time1=="Grassland")

```

##  Plot the change in basal area by severity class
Corset plots are a visualization technique used to visualize repeat measures at 2 time points (such as pre- and post- data). Specifically, corset plots visualize the distribution of measurements at each time point, as well as the trajectories of individual change. This method of visualization is ideal for showing the heterogeneity of data and differences by sub-groups. See: https://cran.r-project.org/web/packages/ggcorset/vignettes/corset_plot_intro.html

A corset plot consists of:

--distribution at each time point using half violins (overall distributions or distributions by group);

--individual trajectories of change between each time point; and optionally

--‘eyelets’ showing either the standard error mean (SEM), or (option used here) mean +/- 1 standard deviation (SD) by group

Note: The faceting option in gg_corset only allows one grouping and I haven't found a work around for that issue yet. Just to demonstrate what the plots look like, I've subset the data to include one cover type in each plot with four levels of severity.

```{r bachange, echo=FALSE, warning=FALSE, fig.height = 9, fig.width = 9, fig.align = "center"}

p1<-gg_corset(ct1, y_var1 = "BA.time1", y_var2 = "BA.time2", 
  group = "Plot_ID", c_var = "severity.time1", eyelets = T, faceted=T, line_size=0.5, e_type="SD") +
  theme_ggcorset() + 
  ggtitle("Ponderosa Pine Forest: Change in Basal Area") + 
  ylab("Basal Area") + xlab("") +
  scale_x_discrete(labels = c("Time 1","Time 2")) +
  guides(colour = guide_legend(override.aes = list(size = 3))) + 
  theme(legend.title=element_blank())
p2<-gg_corset(ct2, y_var1 = "BA.time1", y_var2 = "BA.time2", 
  group = "Plot_ID", c_var = "severity.time1", eyelets = T, faceted=T, line_size=0.5, e_type="SD") +
  theme_ggcorset() + 
  ggtitle("Mixed Conifer Forest: Change in Basal Area") + 
  ylab("Basal Area") + xlab("") +
  scale_x_discrete(labels = c("Time 1","Time 2")) +
  guides(colour = guide_legend(override.aes = list(size = 3))) + 
  theme(legend.title=element_blank())
p3<-gg_corset(ct3, y_var1 = "BA.time1", y_var2 = "BA.time2", 
  group = "Plot_ID", c_var = "severity.time1", eyelets = T, faceted=T, line_size=0.5, e_type="SD") +
  theme_ggcorset() + 
  ggtitle("Conifer Woodland: Change in Basal Area") + 
  ylab("Basal Area") + xlab("") +
  scale_x_discrete(labels = c("Time 1","Time 2")) +
  guides(colour = guide_legend(override.aes = list(size = 3))) + 
  theme(legend.title=element_blank())
p4<-gg_corset(ct4, y_var1 = "BA.time1", y_var2 = "BA.time2", 
  group = "Plot_ID", c_var = "severity.time1", eyelets = T, faceted=T, line_size=0.5, e_type="SD") +
  theme_ggcorset() + 
  ggtitle("Broadleaf Evergreen Woodland: Change in Basal Area") + 
  ylab("Basal Area") + xlab("") +
  scale_x_discrete(labels = c("Time 1","Time 2")) +
  guides(colour = guide_legend(override.aes = list(size = 3))) + 
  theme(legend.title=element_blank())
p5<-gg_corset(ct5, y_var1 = "BA.time1", y_var2 = "BA.time2", 
  group = "Plot_ID", c_var = "severity.time1", eyelets = T, faceted=T, line_size=0.5, e_type="SD") +
  theme_ggcorset() + 
  ggtitle("Transition Forest: Change in Basal Area") + 
  ylab("Basal Area") + xlab("") +
  scale_x_discrete(labels = c("Time 1","Time 2")) +
  guides(colour = guide_legend(override.aes = list(size = 3))) + 
  theme(legend.title=element_blank())
p6<-gg_corset(ct6, y_var1 = "BA.time1", y_var2 = "BA.time2", 
  group = "Plot_ID", c_var = "severity.time1", eyelets = T, faceted=T, line_size=0.5, e_type="SD") +
  theme_ggcorset() + 
  ggtitle("Grassland: Change in Basal Area") + 
  ylab("Basal Area") + xlab("") +
  scale_x_discrete(labels = c("Time 1","Time 2")) +
  guides(colour = guide_legend(override.aes = list(size = 3))) + 
  theme(legend.title=element_blank())

```

```{r patch, echo=FALSE, warning=FALSE, fig.height = 9, fig.width = 9, fig.align = "center"}
## no real benefit to this so plot each separately
## p1 + p2 + p3 + p4 + p5 + p6
p1
p2
p3
p4
p5
p6


```

