---
title: "terrain plots"
author: "S. Haire"
date: "`r Sys.Date()`"
output: 
  html_document: 
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 

library(ggplot2)
library(tidyr)
library(dplyr)
library(ggpubr)
library(readr)
library(patchwork)
library(wesanderson)
library(randomcoloR)

# number of islands
newpal5<- distinctColorPalette(5)
## number of comm classes
newpal8<- distinctColorPalette(8)
pal6=wes_palette("GrandBudapest1")
newpal2<-c('#191970','#ff0000',
'#ffd700','#00ff00','#00ffff','#0000ff',
'#ffb6c1','#006400') #https://mokole.com/palette.html
```

```{r dataprep, include=FALSE}
## read data
d<-read_csv("./data/allvars_v1.csv")
names(d)
## sample size by covertype (print table in Rmd)
d %>% group_by(covertype) %>% count() %>% arrange(-n)
## change index
d$delta.ba<-d$BA.time2 - d$BA.time1
d$delta.tph<-d$TPH.time2 - d$TPH.time1
types<-c("Mixed Conifer Forest", "Broadleaf Evergreen Woodland",
                     "Ponderosa Pine Forest", "Conifer Woodland")
fwdat<- d %>% filter(covertype %in% types)
fwdat$coverf<-factor(fwdat$covertype, levels=c("Mixed Conifer Forest", "Broadleaf Evergreen Woodland", "Ponderosa Pine Forest", "Conifer Woodland"))
## circ plot: aspect and slope
fwdat$aspect.deg<-fwdat$aspect*180/pi
fwdat$slope.deg<-fwdat$slope*180/pi

```

### Exploring terrain metrics at the field plot locations

#### Topography can influence many ecological processes because it is closely related to microclimates and their effects on soil properties, seed dispersal and germination, tree regeneration and growth. It is also a major factor in fire spread, both in facilitating spread and creating shelters or firebreaks. This report takes a first look at the topographic characteristics of the plot locations by covertype and by mountain range. 

#### First, a summary of sample size for each cover type.

```{r covertypes, echo=FALSE}
fwdat %>% group_by(covertype) %>% count() %>% arrange(-n)
```

### Plots of topographic variables

```{r terrainplots, echo=FALSE}
p1<-ggplot(fwdat, aes(elevation, fill=covertype, color=covertype)) +
geom_histogram(aes(y = ..count..), binwidth = 50) +
  scale_fill_manual(values=c("#DFB968", "#D1CACA", "#7AD4D5", "#C059DE")) + 
  theme(legend.position="left",legend.title=element_blank()) +
  xlab("Elevation (m)") + ggtitle("Elevation") +
  facet_wrap(vars(isl_fld), nrow=1)

p2<-ggplot(fwdat, aes(elevation, fill=covertype, color=covertype)) +
geom_histogram(aes(y = ..count..), binwidth = 50) +
  scale_fill_manual(values=c("#DFB968", "#D1CACA", "#7AD4D5", "#C059DE")) + 
  theme(legend.position="none",legend.title=element_blank()) +
  xlab("Elevation (m)") + ggtitle("") 

p3<-ggplot(fwdat, aes(heat.load.index, fill=covertype, color=covertype)) +
geom_histogram(aes(y = ..count..), binwidth = 0.1) +
  scale_fill_manual(values=c("#DFB968", "#D1CACA", "#7AD4D5", "#C059DE")) + 
  theme(legend.position="left",legend.title=element_blank()) +
  xlab("HLI") + ggtitle("Heat Load Index") +
  facet_wrap(vars(isl_fld), nrow=1)

p4<-ggplot(fwdat, aes(heat.load.index, fill=covertype), color=covertype) +
geom_histogram(aes(y = ..count..), binwidth = 0.1) +
  scale_fill_manual(values=c("#DFB968", "#D1CACA", "#7AD4D5", "#C059DE")) + 
  theme(legend.position="none",legend.title=element_blank()) +
  xlab("HLI") + ggtitle("")

p5<-ggplot(fwdat, aes(relative.position, fill=covertype, color=covertype)) +
geom_histogram(aes(y = ..count..), binwidth = 0.1) +
  scale_fill_manual(values=c("#DFB968", "#D1CACA", "#7AD4D5", "#C059DE")) + 
  theme(legend.position="left",legend.title=element_blank()) +
  xlab("RP") + ggtitle("Relative Position") +
  facet_wrap(vars(isl_fld), nrow=1)

p6<-ggplot(fwdat, aes(relative.position, fill=covertype, color=covertype)) +
geom_histogram(aes(y = ..count..), binwidth = 0.1) +
  scale_fill_manual(values=c("#DFB968", "#D1CACA", "#7AD4D5", "#C059DE")) + 
  theme(legend.position="none",legend.title=element_blank()) +
  xlab("RP") + ggtitle("") 

p7<-ggplot(fwdat, aes(topo.ruggedness.index, fill=covertype, color=covertype)) +
geom_histogram(aes(y = ..count..), binwidth = 1) +
  scale_fill_manual(values=c("#DFB968", "#D1CACA", "#7AD4D5", "#C059DE")) + theme(legend.position="right",legend.title=element_blank()) +
  xlab("TRI") + ggtitle("Topographic Ruggedness Index") +
  facet_wrap(vars(isl_fld), nrow=1)

p8<-ggplot(fwdat, aes(topo.ruggedness.index, fill=covertype, color=covertype)) +
geom_histogram(aes(y = ..count..), binwidth = 1) +
  scale_fill_manual(values=c("#DFB968", "#D1CACA", "#7AD4D5", "#C059DE")) + 
  theme(legend.position="none",legend.title=element_blank()) +
  xlab("TRI") + ggtitle("") 

p13<-ggplot(fwdat, aes(topo.wetness.index, fill=covertype, color=covertype)) +
geom_histogram(aes(y = ..count..), binwidth = 1) +
  scale_fill_manual(values=c("#DFB968", "#D1CACA", "#7AD4D5", "#C059DE")) + theme(legend.position="right",legend.title=element_blank()) +
  xlab("TWI") + ggtitle("Topographic Wetness Index") +
  facet_wrap(vars(isl_fld), nrow=1)

p14<-ggplot(fwdat, aes(topo.wetness.index, fill=covertype, color=covertype)) +
geom_histogram(aes(y = ..count..), binwidth = 1) +
  scale_fill_manual(values=c("#DFB968", "#D1CACA", "#7AD4D5", "#C059DE")) + 
  theme(legend.position="none",legend.title=element_blank()) +
  xlab("TWI") + ggtitle("") 

p9<-ggplot(fwdat, aes(aspect.deg, fill=covertype, color=covertype)) +
geom_histogram(aes(y = ..count..), binwidth = 30) +
  scale_x_continuous(limits = c(0, 360)) +#, breaks=sequence(0,360, by=2)) + 
  scale_fill_manual(values=c("#DFB968", "#D1CACA", "#7AD4D5", "#C059DE")) + 
  theme(legend.title=element_blank()) +
  xlab("") + ggtitle("Aspect") +
coord_polar(theta = "x", start = 0, direction = 1) +
  facet_wrap(vars(isl_fld), nrow=1)

p10<-ggplot(fwdat, aes(aspect.deg, fill=covertype, color=covertype)) +
geom_histogram(aes(y = ..count..), binwidth = 30) +
  scale_x_continuous(limits = c(0, 360)) +#, breaks=sequence(0,360, by=2)) + 
  scale_fill_manual(values=c("#DFB968", "#D1CACA", "#7AD4D5", "#C059DE")) + 
  theme(legend.position="left",legend.title=element_blank()) +
  xlab("") + ggtitle("") +
coord_polar(theta = "x", start = 0, direction = 1) 

p11<-ggplot(fwdat, aes(slope.deg, fill=covertype, color=covertype)) +
geom_histogram(aes(y = ..count..), binwidth = 5) +
  scale_x_continuous(limits = c(0, 50)) +
  scale_fill_manual(values=c("#DFB968", "#D1CACA", "#7AD4D5", "#C059DE")) + 
  theme(legend.title=element_blank()) +
  xlab("") + ggtitle("Slope") +
coord_polar(theta = "x", start = 0, direction = 1) +
  facet_wrap(vars(isl_fld), nrow=1)

p12<-ggplot(fwdat, aes(slope.deg, fill=covertype, color=covertype)) +
geom_histogram(aes(y = ..count..), binwidth = 5) +
  scale_x_continuous(limits = c(0, 50)) +
  scale_fill_manual(values=c("#DFB968", "#D1CACA", "#7AD4D5", "#C059DE")) + 
  theme(legend.title=element_blank()) +
  xlab("") + ggtitle("") +
coord_polar(theta = "x", start = 0, direction = 1) 

```

```{r plotlayouts, echo=FALSE, fig.height=8, fig.width=11, fig.align='center'}
p1/p2 + plot_layout(guides='collect')
p3/p4 + plot_layout(guides='collect')
p5/p6 + plot_layout(guides='collect')
p7/p8 + plot_layout(guides='collect')
p13/p14 + plot_layout(guides='collect')
p9/p10 + plot_layout(guides='collect', widths=c(3,2))
p11 /p12 + plot_layout(guides='collect', widths=c(3,2))

```


